---
title: "Modeling Tesla Stock Price Using Tweets"
author: "Tilina Alzaben, Jay Chung, Marion Haney, Divya Rao"
date: "04-21-2024"
documentclass: article
geometry: margin=1in
output:
  bookdown::pdf_document2:
    toc: no
    includes:  
      in_header: preamble-latex.tex
---

```{r, include=FALSE}
###########################
# STYLE EDITS: IGNORE THIS
###########################

# normally you'll want to include this with the libraries at the beginning of your document
knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings
knitr::opts_chunk$set(echo = FALSE) # set echo = FALSE to hide code from output
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(tidytext))
suppressMessages(library(textdata))
suppressMessages(library(dplyr))
suppressMessages(library(quantmod))
suppressMessages(library(fGarch))
```

# Executive summary {-}

...

# Introduction

Public opinion of a company can be beneficial or harmful to a company's 
image and may also affect their sales or profits. Thus, every large company 
hires employees to benefit their branding, marketing, and public communications. 
We set to investigate if we can quantify a relationship between a company's 
public image and their financial health using the case study of Tesla.

Tesla, a company which specializes in electric vehicles, is currently led by 
Elon Musk, who is often intertwined with the company's public image. Musk can
be considered a public relations nightmare, engaging in Twitter rants and having
a lack of filter that often leads to controversies and speculation.

To measure public opinion, we use textual data gathered from Twitter, a 
text-based social media with millions of English-speaking users. Tweets 
mentioning Tesla or the company's stock ticker, TSLA, over one year were 
collected (09-30-2021 to 09-30-2022). From the tweets, we assign sentiment 
scores based on the positive and negative language used from 
the text of the tweets, giving us a measurable proxy for public opinion. 
To measure the financial health of Tesla, we observe its adjusted closing 
stock price over same the year. 

We chose Tesla because of our background knowledge of its popularity within the 
public eye, giving us a wealth of positive and negative examples for sentiment 
analysis. Also, Tesla's stock TSLA is a notably volatile stock, jumping between 
$\209.38$ and $\$409.97$ several times during the year of data.

We engage in time series analysis and attempt to u


## Research Questions
- Does sentiment of Tesla-related tweets have a relationship with Tesla’s stock returns on a given day?
- Can we use the sentiment of tweets mentioning Tesla to forecast Tesla’s stock returns on a given day?

## Data
For this analysis, we utilized two related datasets from Kaggle.  

Twitter tweets: 80793 tweets, 25 companies (including Tesla)

Stock Market Data: Daily close price of stock for the same 25 companies

Companies are the top 25 “most watched” stocks from Yahoo Finance and we 
filtered our data to only include tweets that mention Tesla or TSLA stock.
This left us with 46% of the tweets in data mentioned TSLA (37422 tweets) 
across the year of data collection.

Date range: 09-30-2021 to 09-30-2022 (excluding weekends).


...

# Methods

## Exploratory Data Analysis

Exploratory Data Analysis (EDA) was conducted for deciding on reasonable 
data transformations and analysis, determining which type of model to use, and 
choosing the parameters for these models. 

```{r}
# Gather stock tweets data
tweets <- read.csv("/Users/marionhaney/Library/Mobile Documents/com~apple~CloudDocs/CMU/36671 Time Series/time-series-tweets-stocks/data/stock_tweets.csv")
# Retaining the year, month, day
tweets$day <- as.Date(tweets$Date,"%Y-%m-%d %H:%M:%S")
# Filter to just Tesla tweets
tesla <- filter(tweets, tweets$Company.Name == "Tesla, Inc.")
num_tweets_tesla <- data.frame(table(tesla$day))
names(num_tweets_tesla) <- c("day", "num_tweets")
# Time series for number of Tesla tweets per day
tesla_ts <- ts(num_tweets_tesla$num_tweets, 
                    start = c(2021, 273),
                    frequency = 365)
```

## Number of Tesla Tweets Per Day

Our data included tweets from 09-30-2021 to 09-30-2022 (excluding weekends) 
which mentioned either Tesla's stock ticker or the company name itself. Our 
data included 37,422 tweets regarding Tesla over the year. Here, we view the
number of tweets mentioning Tesla per day plotted over time
along with the ACF and PACF plots of the time series.

```{r}
plot(tesla_ts, main = "Number of Tesla Tweets \n Per Day",
     ylab = "Number of Tweets")
```

```{r}
par(mfrow=c(1,2))
acf(tesla_ts, main = "ACF of Number of Tesla Tweets")
pacf(tesla_ts, main = "PACF of Number of Tesla Tweets")
```

The time series of the number of Tesla tweets per day shows non-stationary 
characteristics, with noticeable fluctuations and periodic spikes indicating 
changes in the mean and variance over time. The ACF plot specifically shows 
possible seasonality with four clusters of ACF spikes over the time period of 
the data. These spikes in the number of tweets mentioning Tesla appear to 
occur approximately every three months.

Since the number of Tesla tweets varies per day in a periodic manner, we 
decided to use averaging in our calculation of daily tweet sentiment. Thus, 
the fluctuating number of Tesla tweets is accounted for.


## Daily Sentiment of Tesla Tweets

For our analysis, we calculated a daily score representing the average 
sentiment (positive or negative) for Tesla tweets posted on the given day.
Our approach involved breaking each tweet into a bag of words 
representation, where each word is given a sentiment score of 1 if it is 
positive, -1 if it is negative, and 0 otherwise (neutral). These individual
word scores were then summed into a sentiment score for each tweet. The 
sentiment scores per tweet ranged from -9 to 9. Finally, the individual 
tweet sentiment scores were averaged into a daily tweet sentiment score for the 
day. The daily sentiment scores ranged from -0.575 to 1.65.

```{r}
df <- read.csv("/Users/marionhaney/Library/Mobile Documents/com~apple~CloudDocs/CMU/36671 Time Series/time-series-tweets-stocks/data/stock_tweets.csv")  %>%
  filter(Stock.Name == 'TSLA') %>%
  mutate(Tweet.ID = row_number()) %>%
  dplyr::select(Tweet.ID, Date, Tweet)
#dim(df)
#names(df)

df$Date <-  ymd(substr(df$Date, 1, 10))
tweets <- data.frame(df)

```

```{r}
# Sentiment analysis
map_bing_sentiment <- function(sentiment) {
  ifelse(sentiment %in% c("positive"), 1, ifelse(sentiment %in% c("negative"), -1, 0))
}

map_nrc_sentiment <- function(sentiment) {
  nrc_positive_sentiments <- c("positive", "anticipation", "surprise", "trust", "joy")
  nrc_negative_sentiments <- c("negative", "anger", "disgust", "fear", "sadness")
  ifelse(sentiment %in% nrc_positive_sentiments, 1,
         ifelse(sentiment %in% nrc_negative_sentiments, -1, 0))
}

tweet_tokens <- tweets %>%
  unnest_tokens(word, Tweet)  

sentiments <- get_sentiments("bing") %>% mutate(sentiment_score = map_bing_sentiment(sentiment))

tweets_sentiment <- tweet_tokens %>%
  inner_join(sentiments, by = "word", relationship = "many-to-many")  %>%
  distinct(Tweet.ID, Date, word, .keep_all = TRUE)

tweets_sentiment_summary <- tweets_sentiment %>%
  group_by(Tweet.ID, Date) %>%  
  summarise(sentiment_score = sum(sentiment_score, na.rm = TRUE), .groups = "drop")

daily_sentiment <- tweets_sentiment_summary  %>%
  group_by(Date) %>%
  summarise(daily_sentiment = mean(sentiment_score))
```

Here, we can view the daily sentiment of tweets over the year of our data 
along with the ACF and PACF.
```{r}
sentiment_ts <- ts(daily_sentiment$daily_sentiment, 
                    start = c(2021, 273),
                    frequency = 365)
plot(sentiment_ts, main = "Daily Sentiment of Tesla Tweets",
     ylab = "Average Daily Sentiment Score")
```

```{r}
par(mfrow=c(1,2))
acf(sentiment_ts, main = "ACF of Daily Sentiment")
pacf(sentiment_ts, main = "PACF of Daily Sentiment")
```

Similarly to the number of Tesla tweets per day, we observe non-stationary 
characteristics for the daily sentiment of Tesla tweets. we see four sections of
ACF spikes approximately 3 months apart, suggesting seasonality. However, the 
magnitude of the spikes for daily sentiment of tweets is less severe than the 
magnitude of the corresponding spikes for the number of tweets per day. This is
potentially due to the averaging of each individual tweet's sentiment score to 
produce a daily metric.


## TSLA Stock Price and Returns

We calculated the returns of the TSLA stock by differencing the adjusted 
close price of TSLA to 1 degree.

```{r}
df <- read.csv("/Users/marionhaney/Library/Mobile Documents/com~apple~CloudDocs/CMU/36671 Time Series/time-series-tweets-stocks/data/stock_yfinance_data.csv") %>%
  filter(Stock.Name == 'TSLA') %>%
  dplyr::select(Date, Adj.Close)
df$Date <- as.Date(df$Date)
df$Returns <- c(diff(df$Adj.Close), NA)
stock_ts <- ts(df$Returns, 
                    start = c(2021, 273),
                    frequency = 365)
df$Returns <- c(diff(log(df$Adj.Close)), NA)
stock_ts_lsq <- ts(df$Returns^2, 
                    start = c(2021, 273),
                    frequency = 365)
stock <- data.frame(df)
stock <- stock %>% na.omit()
```

```{r}
# EDA TSLA returns
plot(stock_ts, main="TSLA Returns")
par(mfrow=c(1,2))
acf(stock_ts, main="ACF of TSLA Returns", na.action = na.pass)
pacf(stock_ts, main="PACF of TSLA Returns", na.action = na.pass)
```

For TSLA returns, we see slight non-stationary characteristics. The ACF and 
PACF plots exhibit spikes outside the white noise error bounds at around 
lag 0.02 and 0.05. The time series plot of TSLA returns also exhibits some 
evidence of heteroskedasticity, with a truncated shape of the time series. The 
earlier returns exhibit a wider variance than the later returns.

Thus, we decided to square the logged returns to alleviate the slight 
non-stationary characteristics and heteroskedasticity.

```{r}
plot(stock_ts_lsq, main="TSLA Squared Log Returns")
par(mfrow=c(1,2))
acf(stock_ts_lsq, main="ACF of TSLA \nSquared Log Returns", na.action = na.pass)
pacf(stock_ts_lsq, main="PACF of TSLA \nSquared Log Returns", na.action = na.pass)
```

The square of logged TSLA returns exhibits only one ACF and PACF spike at 
around lag 0.03 and the plot of the square of logged TSLA returns no longer 
has a truncated shape.

So, we continued with using the square of logged TSLA returns to represent the 
measure of TSLA stock price in our model.

## GARCH Model
Based on the non-constant variance and non-stationarity seen from the 
EDA for daily sentiment, we decided to use a GARCH model. A GARCH 
(generalized autoregressive conditionally heteroskedastic) model uses values 
of the past observations and variances to model the variance at time t. GARCH 
is used commonly with financiald ata because of the high volatility. Since we 
calculated the order of the model from running auto.arima(), we decided to fit a 
GARCH(1, 1) model to returns, the difference of the log of the adjusted 
closing price.

### auto.arima()
The auto.arima() function was applied on the squared logged returns, which is 
the difference of the log of the adjusted closing price squared. This 
helped identify if there was autoregression or moving average pattern 
in the votatility of the returns and helps determine the order of
the GARCH model. 

```{r}
suppressWarnings(library(forecast))

arma_rt_squared <- auto.arima(stock$Returns^2, max.p = 5, max.q = 5, max.order = 10,
                              stationary = T, seasonal = F, trace = F,
                              stepwise = F, approximation = F)
summary(arma_rt_squared)
```

## VAR
We used a VAR (Vector Autoregression) model for feature selection for 
the two linear models. After running the VARselect() function, we found that 
the lagged values of l1 and l2 and lagged values of returns.



# Results

```{r}
stock_ts <- ts(stock$Returns, 
               start=c(2021,9), 
               frequency=365)  
garch_model <- garchFit(~ garch(1,1), data=stock_ts, trace=FALSE)
summary(garch_model)

par(mfrow=c(1,1))
plot(garch_model, which = 1)
plot(garch_model, which = 2)

par(mfrow=c(1,1))
```


```{r}
combined_data <- left_join(stock, daily_sentiment, by = "Date")
#cor(combined_data$daily_sentiment, combined_data$Returns, use = "complete.obs")
model <- lm(Returns ~ daily_sentiment, data = combined_data)
summary(model)
```

```{r}
suppressMessages(library(vars))
df <- combined_data[, c("Returns", "daily_sentiment")]
# VARselect
lag.select <- VARselect(df, 
                        lag.max = 30, 
                        type = "both")
optimal.lags <- lag.select$selection['AIC(n)']

# Fit the VAR model
var.model <- VAR(df, p = optimal.lags)

summary(var.model)
```

```{r}
suppressMessages(library(dplyr))

combined_data <- combined_data %>%
  arrange(Date) %>%
  mutate(
    Returns_l1 = lag(Returns, 1),
    Returns_l2 = lag(Returns, 2),
    daily_sentiment_l1 = lag(daily_sentiment, 1),
    daily_sentiment_l2 = lag(daily_sentiment, 2)
  )

model <- lm(Returns ~ Returns_l1 + Returns_l2 + daily_sentiment + daily_sentiment_l1 + daily_sentiment_l2, data = combined_data)

summary(model)
```

```{r}
model <- lm(daily_sentiment ~ Returns + Returns_l1 + Returns_l2 + daily_sentiment_l1 + daily_sentiment_l2, data = combined_data)

summary(model)
```


# Discussion
Our goal for this project was to predict Tesla's stock price based on the 
sentiment of the tweets. For the first linear model, we are predicting returns
based on past returns, current daily sentiment, past daily sentiment. We see 
only one statistically significant coefficient, which means none of the other 
features have a statistically significant coefficient other than L2 daily 
sentiment. However, L2 daily sentiment has a small magnitude of 0.02, reflecting 
a weak correlation between returns and L2 daily sentiment. In the second linear 
model, we are predicting daily sentiment using returns, past returns, past 
daily sentiment. Almost all of the features are statistically significant value 
other than current returns. The adjusted $R^2$ of 10% suggests that these 
features account for 10% of the variability of daily sentiment. Past returns 
and past daily sentiment Granger causes current daily sentiment, i.e.
past returns and past daily sentiment can be used to predict the 
daily sentiment for the current day. However, our original hypothesis was that 
we would be able to predict returns and stock price using past returns and 
daily sentiment, which is the opposite direction of Granger causality.

The findings from our model indicate a limited capacity for using tweet 
sentiment and frequency to predict Tesla's stock price. The significant but 
minimal correlation found in L2 daily sentiment suggests that while some 
predictive power exists, it is not strong enough to influence financial 
decisions effectively. The ability to predict daily sentiment from past 
stock prices and sentiment highlights potential areas for deeper exploration 
into the dynamics between social media and financial markets.


## Conclusion
TSLA Stock returns cannot be predicted based on the current or lagged 1 values 
of daily sentiment or lagged 1 or 2 values for returns. The coefficient for 
lagged 2 daily sentiment is statistically significant but too small to be 
useful. Daily sentiment cannot be predicted based on current stock returns but 
can be explained by the lagged 1 or 2 values of the stock price or daily 
sentiment. 


## Future Work

As we move forward, our project will concentrate on enhancing the predictive 
models and broadening the analytical scope to better understand the 
relationship between social media activities and stock market behavior. 
Future efforts will include incorporating broader market performance 
indicators to understand Tesla’s stock movements within the larger financial 
ecosystem. We will also explore longer temporal analyses to determine whether 
the effects of negative social media events on stock prices are only temporary 
or if they have a lasting impact. Additionally, we plan to incorporate
tweet engagement metrics such as likes, retweets, and views, which may offer 
deeper insights into how social media dynamics can influence stock price
movements. Through these improvements, we aim to build stronger predictive 
tools for financial market analysis, refining our approach to use the 
predictive power of social media sentiment more effectively.


...

\pagebreak

# Appendix {-}

Notice how the appendix below gathers all the code blocks above and nicely pastes them together.

```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```