# Number of tweets as a time series
Investigating the number of tweets per day as a time series: EDA.

Load libraries:
```{r}
library(tidyverse)
```

Load data:
Tweets mentioning top 25 most watched stock tickers on Yahoo Finance
Start: 9-30-2021
End: 9-30-2022
```{r}
df <- read.csv("data/stock_tweets.csv")
dim(df)
```

Convert the 'Date' column to a DateTime class instead of a string:
```{r}
# Retaining the year, month, day
df$day <- as.Date(df$Date,"%Y-%m-%d %H:%M:%S")
head(df)
```

EDA for number of tweets per day:
```{r}
# Cast 'day' column to a string version to count
df$str_day <- as.character(df$day)

num_tweets_day <- data.frame(table(df$str_day))
names(num_tweets_day) <- c("day", "num_tweets")

head(num_tweets_day)
```

Turn the number of tweets into a time series:
```{r}
num_tweets_ts <- ts(num_tweets_day$num_tweets, 
                    start = c(2021, 273),
                    frequency = 365)
head(num_tweets_ts)
```

Plot time series:
```{r}
plot(num_tweets_ts, 
     main = "Number of Tweets per day\nTweets mentioning top 25 most watched stock tickers on Yahoo Finance",
     ylab = "Number of Stock Tweets")
```

Plot ACF:
```{r}
acf(num_tweets_ts, ylab = "ACF", 
    main = "ACF of Number of Stock Tweets")
```

Plot PACF:
```{r}
pacf(num_tweets_ts, ylab = "PACF", 
     main = "PACF of Number of Stock Tweets")
```

25 companies in data:
```{r}
sort(unique(df$Company.Name))
```



